---
title: "Lesson3"
author: 'noman'
output: html_document
---

# Lesson3: R Markdown and Basic Functions

## Read the 'datasets.xlsx'

R can not read an xl dataset directly. So, first install `readxl`, then load it to use its `read.xl()` fuction.

#To install the package or library , consol:install.packages('readxl'), or Tools\>Install Packages, or Packages in the file pane.\

#To insert a command line: click C+. or `ctrl+alt+i`

```{r}
library(readxl)
df=read_excel('datasets.xlsx', sheet = 'wrangling', range = 'B7:N507')
head(df)


```

```{r}
tail(df)

```

#Calculating the column sum, excluding the colm no 1, only for 1 to 50 rows

```{r}
colSums(df[c(1:50),c(2:13)])
```

#To see the full `dataset`, excluding the 1st colm containing the money scarcity information

```{r}
colSums(df[-1])
```

Now give a name for this object and such as scarcity. Then express the frequency as percentage.

```{r}
scarcity=colSums(df[-1])*100/nrow(df)
scarcity
```

#Visualize this month wise scarcity using ggplot [ggplot2 package]. First install ggplot2 package.

Scarcity is not a data frame but ggplot fumnction need data frame(with a few variables). convert scarcity to a data frame df_scarcity.

convert the row names to column named as `month` by using `dplyr`.

```{r}
library(tidyverse)

#convert to data frame
df_scarcity=as.data.frame(scarcity)

#Transfer the row names to column as 'Month'
df_plot=rownames_to_column(df_scarcity, var= 'Month')

#Check the structure and dataframe
str(df_plot)
df_plot
```

Now visualization. To set the order of the month names we need to reorder the month names using

```{r}
df_plot$Month= factor(df_plot$Month, 
                      levels = c('jan', 'feb', 'mar','apr','may', 'jun', 'jul','aug','sep', 'oct', 'nov', 'dec'),
                      labels = c('Jan', 'Feb', 'Mar', 'Apr', 'May', 'Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec'))


```

```{r}
ggplot(df_plot)+
  aes(x=Month, y=scarcity, color=Month, fill=Month)+
  geom_col(width = .7)+
  labs(x='', y= 'Money scarcity')+
  geom_text(aes(label=scarcity,y=scarcity+2), color='grey40')+
  theme_test()+
  theme(legend.position = 'top')

ggsave('my plot.png', dpi=500, height = 4.5, width = 8, units = 'in')

```
